### 2021-07-28 index.Rmd
mayor update repo
- se elimina docs
- paper pasa a estar en el root
- prep data y analysis se mantienen en production

---
author: " "
bibliography:
- 'input/bib/zlibrary.bib'
csl:
- 'input/bib/apa6.csl'
title: " "
output:
    bookdown::html_document2:
        smart: false
        includes:
          in_header: "input/head_github-activity.html"
---

<br>

# This is the project website for the paper "Economic inequality and its influence on the justification of legitimate income differences. {.left}

Last update: `r format(Sys.time(), "%A %d %B %Y %X")`

# Authors:

- Julio Iturra, Centro de Estudios de Conflicto y Cohesión Social - COES [julioiturra@uchile.cl](julioiturra@uchile.cl)
- Juan Carlos Castillo, Departamento de Sociología, Universidad de Chile.
- Catalina Rufs, Pontificia Universidad Católica de Chile.
- Luis Maldonado, Instituto de Sociología, Pontificia Universidad Católica de Chile.

Based on the study "Political and Social Beliefs" conducted in 2014 using a representative sample of the Metropolitan Region of Chile, we replicate the original experimental study conducted by [Trump (2018)](https://doi.org/10.1017/S0007123416000326). We tested how percieved economic inequality affects afects the global and ocupational based salary gaps in a highly unequal context.

# Structure of the project website:

- The tabs follow the logic of the [IPO protocol](https://github.com/juancarloscastillo/ipo) for academic workflows, based on three main components = **I**nput -> **P**roduction -> **O**utput. Here we show only one the main parts of each component: Data $\rightarrow$ Analysis $\rightarrow$ Paper.

* The complete reproducible files are in the [Github repo: https://jciturras.github.io/desigualdad_recom/](https://jciturras.github.io/desigualdad_recom/)

  - _Data_: pre-processing and general descriptives
  - _Analysis_: Balance, Treatment effect analysis (OLS) and annexes
  - _Paper_: final output

## Reproducibility and documentation

The repository of the project in Github is structured according to the [IPO reproducibility protocol](https://juancarloscastillo.github.io/ipo/).

<br>

## [Github activity feed](https://github.com/jciturras/github-activity)

<div id="feed"></div>

<br>